msAggr_seurat-v1

### 2021-06-16 msAggr_AnalysisCode/msAggr_seurat-v1.Rmd
start work on SCTransform method

```{r setup}
knitr::opts_knit$set(root.dir = "~/Desktop/10XGenomicsData/msAggr_scRNASeq/")
library(dplyr)
library(Seurat)
library(patchwork)
library(ggplot2)
```
store session info
```{r}
sink("msAggr_seurat-v1.20210608")
sessionInfo()
sink()
```

@@

```{r}
projectName <- "msAggr"
```

@@

Add mitochondrial metadata and plot some basic features
```{r}
seurat.object[["percent.mt"]] <- PercentageFeatureSet(seurat.object, pattern = "^mt-")
VlnPlot(seurat.object, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3, pt.size = 0, fill.by = 'orig.ident')
```

@@

```{r}
seurat.object <- JackStraw(seurat.object, num.replicate = 100, dims = 40) #runs ~50 min
seurat.object <- ScoreJackStraw(seurat.object, dims = 1:40)
save.image(paste0(projectName, ".RData"))
```

@@

```{r}
# plot1 <- JackStrawPlot(seurat.object, dims = 1:50)
```
```{r, figures-side, fig.show='hold', out.width="50%"}
ElbowPlot(seurat.object, ndims = 50)
percent.variance(seurat.object@reductions$pca@stdev)

```
Number of PCs describing X% of variance

@@

@@

Not a fan of how "swoopy" it looks. Could be to many dimensions? May as well try 90%

@@

Plot the results
```{r}
DimPlot(seurat.object, 
				reduction = "umap"
				) + ggtitle("msAggr dim24 res0.5")

```{r}
DimPlot(seurat.object,
				reduction = "umap", 
				group.by = "orig.ident"
				) + ggtitle("msAggt dim24 orig.ident")
```

```{r}
saveRDS(seurat.object, file = "msAggr_dim24.rds")
```

24 vs 36 dimensions doesn't seem to make that much difference RE overall relationships. Should probbaly create two separate objects and try a few plotting methods.

>>>>

```{r setup}
knitr::opts_knit$set(root.dir = "~/Desktop/10XGenomicsData/msAggr_scRNASeq/")
library(dplyr)
library(Seurat)
library(patchwork)
library(ggplot2)
library(clustree)
```
store session info
```{r}
sink("msAggr_seurat-v1.20210608")
sessionInfo()
sink()
```

@@

```{r}
projectName <- "msAggr"
jackstraw.dim <- 40
```

@@

Add mitochondrial metadata and plot some basic features
```{r}
seurat.object[["percent.mt"]] <- PercentageFeatureSet(seurat.object, pattern = "^mt-")
VlnPlot(seurat.object, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3, pt.size = 0, fill.by = 'orig.ident', )
```

@@

```{r}
jackstraw.dim <- 40
seurat.object <- JackStraw(seurat.object, num.replicate = 100, dims = jackstraw.dim) #runs ~50 min
seurat.object <- ScoreJackStraw(seurat.object, dims = 1:jackstraw.dim)
save.image(paste0(projectName, ".RData"))
```

@@

```{r}
JackStrawPlot(seurat.object, dims = 25:36)
```
```{r, figures-side, fig.show='hold', out.width="50%"}
ElbowPlot(seurat.object, ndims = 50)
percent.variance(seurat.object@reductions$pca@stdev)
```
Number of PCs describing X% of variance

@@

# KNN dimensional reduction
## Dim36

@@

## Dim24
36 dimensions looks very "swoopy". might have asked for too many dimensions. Will try dim = 24, which accounts for 90% of variance

```{r}
# seurat.object <- FindNeighbors(seurat.object, dims = 1:24)
# seurat.object <- FindClusters(seurat.object, resolution = 0.5)
# seurat.object <- RunUMAP(seurat.object, dims = 1:24)
```

Plot the results
```{r}
# DimPlot(seurat.object, 
# 				reduction = "umap"
# 				) + ggtitle("msAggr dim24 res0.5")

```{r}
# DimPlot(seurat.object,
# 				reduction = "umap", 
# 				group.by = "orig.ident"
# 				) + ggtitle("msAggt dim24 orig.ident")
```

```{r}
# saveRDS(seurat.object, file = "msAggr_dim24.rds")
```

## KNN conclusion
24 vs 36 dimensions doesn't seem to make that much difference RE overall relationships between cells. Should probbaly create two separate objects and try a few plotting methods, but will start by focusing on 36 dimensions.

# Cluster analysis

Will proceed with dim = 36 and do clustering analysis with a range of clusters. Later can do the tree-based over-clustering assessment

```{r}
seurat.object <- readRDS("msAggr_dim36.rds")
```

## Generate cluster profiles
Picking range of resolutions
```{r}
for(x in c(0.5, 1, 1.5, 2, 2.5)){
	seurat.object <- FindClusters(seurat.object, resolution = x)
}
seurat.object <- RunUMAP(seurat.object, dims = 1:36)
```

Plot the clustering over different resolutions. Going to need a much better color palette, or try mixing in some different symbols

```{r}
for (meta.col in colnames(seurat.object@meta.data)){
	if(grepl(pattern = ("RNA_snn_res"), x = meta.col)==TRUE){
		myplot <- DimPlot(seurat.object, 
											group.by = meta.col,
											reduction = "umap", 
											cols = colorRamps::primary.colors(n = length(levels(seurat.object@meta.data[[meta.col]])))
											) + 
			ggtitle(paste0("msAggr dim36 res", gsub("RNA_snn_res", "", meta.col) ))
		plot(myplot)
	}
}
```

```{r}
# saveRDS(seurat.object, file = "msAggr_dim36.rds")
```

## Evaluate cluster stability
Must ensure we have the right cluster stability, that is, cells that start in the same cluster tend to stay in the same cluster. If your data is over-clustered, cells will bounce between groups.

Following [this tutorial by Matt O.].https://towardsdatascience.com/10-tips-for-choosing-the-optimal-number-of-clusters-277e93d72d92.

### Clustree
Previously my favourite has been Clustree, which gives a nice visual
NB: For some reason `clustree::clustree()` didn't work, whereas `library(clustree)` followed by `clustree()` did.

```{r fig.height=5}
clustree(seurat.object, prefix = "RNA_snn_res.", node_colour = "sc3_stability") + 
	scale_color_continuous(low = 'red3', high = 'white')
```

```{r fig.height=5}
clustree(seurat.object, prefix = "RNA_snn_res.", expres = 'data', node_colour = "sc3_stability") + 
	scale_color_continuous(low = 'red3', high = 'white')
```

```{r fig.height=5}
clustree(seurat.object, prefix = "RNA_snn_res.", expres = 'scale.data', node_colour = "sc3_stability") + 
	scale_color_continuous(low = 'red3', high = 'white')
```

```{r fig.height=5}
clustree(seurat.object, prefix = "RNA_snn_res.", expres = 'counts', node_colour = "sc3_stability") + 
	scale_color_continuous(low = 'red3', high = 'white')
```

These data suggest that node stability is aweful! Need to figure out if this is a dimensional reduction error or a clustering error.

# Why are clusters so unstable?
Differences could include:
* cells in each population (cellranger v6 includes more cells than cellranger v1, especially in MEP)
* dimensionality is incorrect
* ScaleData didnt account for regression factors (e.g., "nCounts_RNA" or "nFeatures_RNA")
* Incorrect normalization/scaling method
* Clustering is too strict or not strict enough
* neighborhood analysis used wrong parameters
* Should include mitoC filter (there's a chunk of MEP w/ mitoC @ ~40%)
* SCTransform accounts better for sources of variability

```{r}
# Number of filtered cells left in each pop
sapply(c("LSKm2", "CMPm2", "MEPm", "GMPm"), function(x) (c(nrow(seurat.object@meta.data[seurat.object@meta.data$orig.ident == x,]))))
```

```{r}
for (x in c("LSKm2", "CMPm2", "MEPm", "GMPm")){
	h = hist(seurat.object@meta.data[seurat.object@meta.data$orig.ident == x, 'percent.mt'], breaks = 30, plot = FALSE)
	h$density = h$counts/sum(h$counts)*100
	plot(h,freq=FALSE, main =  paste(x, "percent mitoC"), xlab = "percent mitoC", ylab = "Frequency")
}
```

Looks like MEPm is the only sample with that huge MitoC % lump @ 40%. What do these cells look like, otherwise?

```{r fig.width=5}
VlnPlot(subset(seurat.object, subset = orig.ident == "MEPm"), 
				features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 1, pt.size = 0, fill.by = 'ident', flip = TRUE)
```
Save dim36 as is and try clustering analysis @ dim24
```{r}
saveRDS(seurat.object, file = "msAggr_AnalysisCode/msAggr_dim36.rds")
```

# Repeat clustering with dim24
One possibility is that I've included too many dimensions. Will see if 90% increases stability.

@@

Save object
```{r}
saveRDS(seurat.object, file = "msAggr_dim24.rds")
```

```{r}
for (meta.col in colnames(seurat.object@meta.data)){
	if(grepl(pattern = ("RNA_snn_res"), x = meta.col)==TRUE){
		myplot <- DimPlot(seurat.object, 
											group.by = meta.col,
											reduction = "umap", 
											cols = colorRamps::primary.colors(n = length(levels(seurat.object@meta.data[[meta.col]])))
											) + 
			ggtitle(paste0("msAggr dim36 res", gsub("RNA_snn_res", "", meta.col) ))
		plot(myplot)
	}
}
```

## Evaluate cluster stability
Must ensure we have the right cluster stability, that is, cells that start in the same cluster tend to stay in the same cluster. If your data is over-clustered, cells will bounce between groups.

Following [this tutorial by Matt O.].https://towardsdatascience.com/10-tips-for-choosing-the-optimal-number-of-clusters-277e93d72d92.

### Clustree
Previously my favourite has been Clustree, which gives a nice visual
NB: For some reason `clustree::clustree()` didn't work, whereas `library(clustree)` followed by `clustree()` did.

```{r fig.height=5}
clustree(seurat.object, prefix = "RNA_snn_res.", node_colour = "sc3_stability") + 
	scale_color_continuous(low = 'red3', high = 'white')
```

Think I'll explore regression factors using SCTransform in new document.
